---
title: "Project 7: Difference-in-Differences and Synthetic Control"
output: html_document
---


```{r}
# Install and load packages 
if (!require("pacman")) install.packages("pacman")

if (!require("augsynth")) devtools::install_github("ebenmichael/augsynth")

pacman::p_load(# Tidyverse packages including dplyr and ggplot2 
               tidyverse,
               ggthemes,
               augsynth,
               gsynth,
               directlabels)

# set seed
set.seed(1)

# load data
medicaid_expansion <- read_csv('medicaid_expansion.csv')
```

# Introduction

For this project, you will explore the question of whether the Affordable Care Act increased health insurance coverage (or conversely, decreased the number of people who are uninsured). The ACA was passed in March 2010, but several of its provisions were phased in over a few years. The ACA instituted the "individual mandate" which required that all Americans must carry health insurance, or else suffer a tax penalty. There are four mechanisms for how the ACA aims to reduce the uninsured population:

- Require companies with more than 50 employees to provide health insurance.
- Build state-run healthcare markets ("exchanges") for individuals to purchase health insurance.
- Provide subsidies to middle income individuals and families who do not qualify for employer based coverage.
- Expand Medicaid to require that states grant eligibility to all citizens and legal residents earning up to 138\% of the federal poverty line. The federal government would initially pay 100\% of the costs of this expansion, and over a period of 5 years the burden would shift so the federal government would pay 90\% and the states would pay 10\%.

In 2012, the Supreme Court heard the landmark case NFIB v. Sebelius, which principally challenged the constitutionality of the law under the theory that Congress could not institute an individual mandate. The Supreme Court ultimately upheld the individual mandate under Congress's taxation power, but struck down the requirement that states must expand Medicaid as impermissible subordination of the states to the federal government. Subsequently, several states refused to expand Medicaid when the program began on January 1, 2014. This refusal created the "Medicaid coverage gap" where there are indivudals who earn too much to qualify for Medicaid under the old standards, but too little to qualify for the ACA subsidies targeted at middle-income individuals.

States that refused to expand Medicaid principally cited the cost as the primary factor. Critics pointed out however, that the decision not to expand primarily broke down along partisan lines. In the years since the initial expansion, several states have opted into the program, either because of a change in the governing party, or because voters directly approved expansion via a ballot initiative.

You will explore the question of whether Medicaid expansion reduced the uninsured population in the U.S. in the 7 years since it went into effect. To address this question, you will use difference-in-differences estimation, and synthetic control.

# Data

The dataset you will work with has been assembled from a few different sources about Medicaid. The key variables are:

- **State**: Full name of state
- **Medicaid Expansion Adoption**: Date that the state adopted the Medicaid expansion, if it did so.
- **Year**: Year of observation.
- **Uninsured rate**: State uninsured rate in that year.

# Exploratory Data Analysis

Create plots and provide 1-2 sentence analyses to answer the following questions:

- Which states had the highest uninsured rates prior to 2014? The lowest?
- Which states were home to most uninsured Americans prior to 2014? How about in 2021? **Note**: 2010 state population is provided as a variable to answer this question. In an actual study you would likely use population estimates over time, but to simplify you can assume these numbers stay about the same.

```{r}
# highest and lowest uninsured rates

top5 <- medicaid_expansion %>% filter(year == 2014) %>% arrange(uninsured_rate) %>% top_n(c(5)) %>% select(State)

bot5 <- medicaid_expansion %>% filter(year == 2014) %>% arrange(uninsured_rate) %>% top_n(c(-5)) %>% select(State)

p <- medicaid_expansion %>% 
  filter(year <= 2014) %>%
  filter(State %in% c(unlist(top5), unlist(bot5))) %>%
  ggplot(., aes(x = year, 
                y = uninsured_rate, 
                color = State)) + 
  geom_line() +
  scale_colour_discrete(guide = 'none') + 
  ggtitle('5 Highest and 5 Lowest Uninsured Rates by State, 2014') +
  xlab('Year') +
  ylab('Uninsured Rate')  + expand_limits(x = 2015)

direct.label(p, method = "last.qp")





```

```{r}
# most uninsured Americans

top5 <- medicaid_expansion %>% 
  filter(year == 2014) %>% 
  mutate(num_uninsured = uninsured_rate * population) %>%
  arrange(num_uninsured) %>% 
  top_n(c(5)) %>% 
  select(State)

bot5 <- medicaid_expansion %>% 
  filter(year == 2014) %>% 
  mutate(num_uninsured = uninsured_rate * population) %>%
  arrange(num_uninsured) %>% 
  top_n(c(-5)) %>% 
  select(State)

p <- medicaid_expansion %>% 
  filter(year <= 2014) %>%
  filter(State %in% unlist(top5)) %>%
  mutate(num_uninsured = uninsured_rate * population) %>%
  ggplot(., aes(x = year, 
                y = num_uninsured, 
                color = State)) + 
  geom_line() +
  scale_colour_discrete(guide = 'none') + 
  ggtitle('5 Most Uninsured Individuals by State, 2014') +
  xlab('Year') +
  ylab('Uninsured Number')  + expand_limits(x = 2015)

direct.label(p, method = "last.qp")


```

```{r}

p <- medicaid_expansion %>% 
  filter(year <= 2014) %>%
  filter(State %in% unlist(bot5)) %>%
  mutate(num_uninsured = uninsured_rate * population) %>%
  ggplot(., aes(x = year, 
                y = num_uninsured, 
                color = State)) + 
  geom_line() +
  scale_colour_discrete(guide = 'none') + 
  ggtitle('5 Least Uninsured Individuals by State, 2014') +
  xlab('Year') +
  ylab('Uninsured Number')  + expand_limits(x = 2015)

direct.label(p, method = "last.qp")



```


```{r}
medicaid_expansion %>% 
  filter(year == 2020) %>% 
  mutate(num_uninsured = uninsured_rate * population) %>%
  arrange(-num_uninsured) %>% 
  top_n(c(5)) 

```


# Difference-in-Differences Estimation

## Estimate Model

Do the following:

- Choose a state that adopted the Medicaid expansion on January 1, 2014 and a state that did not. **Hint**: Do not pick Massachusetts as it passed a universal healthcare law in 2006, and also avoid picking a state that adopted the Medicaid expansion between 2014 and 2015.
- Assess the parallel trends assumption for your choices using a plot. If you are not satisfied that the assumption has been met, pick another state and try again (but detail the states you tried).

```{r}
# Parallel Trends plot

#California (2014/01/01) and Montana (2016/01/01)

medicaid_expansion %>%
  filter(State %in% c("California","Montana")) %>%
  filter(year >= 2008 & year <= 2020) %>%
  ggplot() + 
  geom_point(aes(x = year, 
                 y = uninsured_rate, 
                 color = State)) +
  geom_line(aes(x = year, 
                y = uninsured_rate, 
                color = State)) +
  geom_vline(aes(xintercept = 2014)) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ggtitle('California and Montana Uninsured Rate \n before/after Medicaid Expansion in 2014') +
  xlab('Year') +
  ylab('State Uninsured Rate')

```

```{r}

#California (2014/01/01) and Virginia (2019/01/01)

medicaid_expansion %>%
  filter(State %in% c("California","Virginia")) %>%
  filter(year >= 2008 & year <= 2020) %>%
  ggplot() + 
  geom_point(aes(x = year, 
                 y = uninsured_rate, 
                 color = State)) +
  geom_line(aes(x = year, 
                y = uninsured_rate, 
                color = State)) +
  geom_vline(aes(xintercept = 2014)) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ggtitle('California and Virginia Uninsured Rate \n before/after Medicaid Expansion in 2014') +
  xlab('Year') +
  ylab('State Uninsured Rate')

```

```{r}

#California (2014/01/01) and Utah (2020/01/01)

medicaid_expansion %>%
  filter(State %in% c("California","Utah")) %>%
  filter(year >= 2008 & year <= 2020) %>%
  ggplot() + 
  geom_point(aes(x = year, 
                 y = uninsured_rate, 
                 color = State)) +
  geom_line(aes(x = year, 
                y = uninsured_rate, 
                color = State)) +
  geom_vline(aes(xintercept = 2014)) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text()) +
  ggtitle('California and Utah Uninsured Rate \n before/after Medicaid Expansion in 2014') +
  xlab('Year') +
  ylab('State Uninsured Rate')

```

- Estimates a difference-in-differences estimate of the effect of the Medicaid expansion on the uninsured share of the population. You may follow the lab example where we estimate the differences in one pre-treatment and one post-treatment period, or take an average of the pre-treatment and post-treatment outcomes

```{r}

# Difference-in-Differences estimation: Using CA and Montana

cm_DnD <- medicaid_expansion %>%
  filter(State %in% c("California","Montana")) %>%
  filter(year >= 2008 & year <= 2020) 
cm_DnD

# FIRST WAY: Difference in one pre-treatment and one post-treatment period

# pre-treatment difference

pre_diff <- cm_DnD %>%
  filter(year == 2013) %>%
  select(State, 
         uninsured_rate) %>%
  spread(State, 
         uninsured_rate) %>%
  summarise(Montana - California)
pre_diff   #-0.00448	

# post-treatment difference

post_diff <- cm_DnD %>%
  filter(year == 2015) %>%
  select(State, 
         uninsured_rate) %>%
  spread(State, 
         uninsured_rate) %>%
  summarise(Montana - California)
post_diff   #0.03014	

# diff-in-diffs

diff_in_diffs <- post_diff - pre_diff
diff_in_diffs   #0.03462	


# SECOND WAY: Difference in average of the pre- and post-treatment outcomes


# pre-treatment difference

pre_diff_avg <- cm_DnD %>%
  filter(year >= 2008 & year <= 2013) %>%
  select(State, year,
         uninsured_rate) %>%
  spread(State, 
         uninsured_rate) %>%
  summarise(mean(Montana) - mean(California)) 
  #Montana avg uninsured rate=0.1811, CA avg = 0.1796

pre_diff_avg  # 0.001460333

# post-treatment difference

post_diff_avg <- cm_DnD %>%
  filter(year >= 2015 & year <= 2020) %>%
  select(State, year,
         uninsured_rate) %>%
  spread(State, 
         uninsured_rate) %>%
  summarise(mean(Montana) - mean(California))
post_diff_avg   # 0.0132822

# diff-in-diffs in average

diff_in_diffs_avg <- post_diff_avg  - pre_diff_avg 
diff_in_diffs_avg   # 0.01182187


```


## Discussion Questions

- Card/Krueger's original piece utilized the fact that towns on either side of the Delaware river are likely to be quite similar to one another in terms of demographics, economics, etc. Why is that intuition harder to replicate with this data?
- **Answer**: The intuition that neighboring towns may be similar to one another is harder to replicate with this data because this data is based on state-level data. States in the U.S. not only are different in demographics and economy, but also have their own governing bodies and are each politically enmeshed in a complicated relationship with the federal government. This difference among states is evident through the different ways in which states responded to the Affordable Care Act and the different timings at which states enacted Medicaid expansion. Therefore using any two neighboring states to measure the effect of Medicaid expansion is not a good comparison to make. 

- What are the strengths and weaknesses of using the parallel trends assumption in difference-in-differences estimates?
- **Answer**: A strength of using the parallel trends assumption in diff-in-diff estimates is that we are able to use an untreated/control unit as a proper counterfactual to the treated unit. In other words, since we cannot know what the treated unit would have been like without the treatment, as long as parallel trends assumption holds, we can take the pre-and post-treatment outcomes of an untreated unit and use that to compare to those of the treated unit to measure the treatment effect. The assumption also allows researchers to make more careful estimations taking into account the passing of time and control for selection bias which one can get from a single difference estimate.  

A weakness of using the parallel trends assumption is that it is a huge assumption to make, especially when the units are clearly not identical. Even if we match the treated and untreated units on baseline covariates, the very fact that the two units are different creates the possibility that their trends may differ over time. And if the parallel trends assumption does not hold, we cannot make a fair comparison to properly measure the treatment effect, since there may be a myriad confounders at play. In addition, the measured treatment effect may vary depending on the time periods observed. As shown above, looking at just one year before and after treatment, and the average of several years before and after treatment produce different diff-in-diff estimates. Depending on how we define the period where the parallel trends assumption holds true, we may be biasing the estimates.


# Synthetic Control

Estimate Synthetic Control

Although several states did not expand Medicaid on January 1, 2014, many did later on. In some cases, a Democratic governor was elected and pushed for a state budget that included the Medicaid expansion, whereas in others voters approved expansion via a ballot initiative. The 2018 election was a watershed moment where several Republican-leaning states elected Democratic governors and approved Medicaid expansion. In cases with a ballot initiative, the state legislature and governor still must implement the results via legislation. For instance, Idaho voters approved a Medicaid expansion in the 2018 election, but it was not implemented in the state budget until late 2019, with enrollment beginning in 2020.

Do the following:

- Choose a state that adopted the Medicaid expansion after January 1, 2014. Construct a non-augmented synthetic control and plot the results (both pre-treatment fit and post-treatment differences). Also report the average ATT and L2 imbalance.

```{r}
# non-augmented synthetic control

#Louisiana (2016-07-01)
# augsynth(outcome ~ trt, unit, time, t_int, data)


#Treatment variable
LA <-  medicaid_expansion %>% 
  mutate(treated = as.numeric(State == "Louisiana" & year >= 2017))

syn <- augsynth(uninsured_rate ~ treated, State, year, 
               LA,
                progfunc = "None", scm = T)

synsum <- summary(syn)

synsum$average_att
synsum$l2_imbalance

```

```{r}

plot(syn)

# this looks decent !

```

- Re-run the same analysis but this time use an augmentation (default choices are Ridge, Matrix Completion, and GSynth). Create the same plot and report the average ATT and L2 imbalance.

```{r}
# augmented synthetic control

rsyn <- augsynth(uninsured_rate ~ treated, State, year, 
               LA,
                progfunc = "Ridge", scm = T)

rsynsum <- summary(rsyn)
rsynsum$average_att
rsynsum$l2_imbalance

plot(rsyn)

```

- Plot barplots to visualize the weights of the donors.

```{r}
# barplots of weights

data.frame(syn$weights) %>%
  # change index to a column
  tibble::rownames_to_column('State') %>%
  ggplot() +
  # stat = identity to take the literal value instead of a count for geom_bar()
  geom_bar(aes(x = State, 
               y = syn.weights),
           stat = 'identity') +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        axis.text.x = element_text(angle = 90)) +
  ggtitle('Synthetic Control Weights') +
  xlab('State') +
  ylab('Weight') 


data.frame(rsyn$weights) %>%
  tibble::rownames_to_column('State') %>%
  ggplot() +
  geom_bar(aes(x = State, y = rsyn.weights),
           stat = 'identity') +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        axis.text.x = element_text(angle = 90)) +
  ggtitle('Synthetic Control Weights') +
  xlab('State') +
  ylab('Weight') 

```

**HINT**: Is there any preprocessing you need to do before you allow the program to automatically find weights for donor states?

## Discussion Questions


- What are the advantages and disadvantages of synthetic control compared to difference-in-differences estimators?

- **Answer**: The advantage with synthetic controls is that they may allow for a more tightly fit counterfactual in the pre-treatment period if otherwise there is not a suitable unit to match Conversely, if there are several well-matched units, they allow for the combination of those data without having to choose one specific unit to match over time. 

A disadvantage could be processing time (depending on if augmentation is used and what method), and more generally, more data is required to be collected in order to create a synthetic control. On a substantive level, there may also be some research questions that are not suitable to answer with synthetic controls as opposed to diff-in-diffs. It may not make sense to think of the counterfactual as a weighted sum of other units. (Even in this case, I'm wondering how to approach why particular donor states are chosen or not chosen.)

- One of the benefits of synthetic control is that the weights are bounded between [0,1] and the weights must sum to 1. Augmentation might relax this assumption by allowing for negative weights. Does this create an interpretation problem, and how should we balance this consideration against the improvements augmentation offers in terms of imbalance in the pre-treatment period?


- **Answer**: I touched on this above, but I think it creates an interpretation problem to the extent that what/which particular units are contributing as donors to the synthetic control is important for the research question. For example, if it's the case that the units are not particularly distinct, then the configuration of units in the synthetic control may not be that meaningful. For this case, however, because the characteristics of the US states tend to be quite different from one another, especially in politics and policies, it may require some more careful interpretation. 

For instance, the top donor for synthetic Louisiana was New Jersey in both regular and augmented estimations. Some of the other donors make more sense (especially those in the South, like Tennessee, Alabama, Texas), but I'm not sure how I would explain why New Jersey. 

# Staggered Adoption Synthetic Control

## Estimate Multisynth

Do the following:

- Estimate a multisynth model that treats each state individually. Choose a fraction of states that you can fit on a plot and examine their treatment effects.


```{r}
#cleaning data - extracting year from Date_Adopted and created a treatment variable to 
#indicate the year a state expands medicaid.

medicaid_expansion <-  medicaid_expansion %>%
  mutate(year_adopted = as.numeric(format(medicaid_expansion$Date_Adopted,'%Y')))%>%
  mutate(year_adopted = ifelse(is.na(year_adopted), 
                               Inf, year_adopted),
         trt = 1 * (year >= year_adopted))
                                   
```



```{r}
# multisynth model states
#mutltisynth(outcome ~ treatment, unit, time, nu, data,  n_leads)

# with a choice of nu
choice_ppool_syn <- multisynth(uninsured_rate ~ trt, State, year, 
                        nu = 0.5, medicaid_expansion , n_leads = 10)


# default nu
default_ppool_syn <- multisynth(uninsured_rate ~ trt, State, year, 
                        medicaid_expansion, n_leads = 10)

print(default_ppool_syn$nu)

choice_ppool_syn
default_ppool_syn

#Both models have the same ATE estimates. Moving forward with the model that uses the default nu.

```

```{r}
ppool_syn_summ <- summary(default_ppool_syn)
ppool_syn_summ
```

```{r}

#plotting the states we examined in the previous sections

ppool_syn_summ$att %>%
  #filter (Level %in% c(unlist(top5), unlist(bot5)))%>%
  filter (Level %in% c("California", "Montana", "Virginia", "Utah", "Louisiana"))%>%
  ggplot(aes(x = Time, y = Estimate, color = Level)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = "bottom") +
  ggtitle('Synthetic Controls for Medicaid Expansion') +
  xlab('Time') +
  ylab('Uninsured Rate')
```
```{r}

# Plotting the states with the 5 least and most uninsured individuals by state in 2014.

ppool_syn_summ$att %>%
  filter (Level %in% c(unlist(top5), unlist(bot5)))%>%
  ggplot(aes(x = Time, y = Estimate, color = Level)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = "bottom") +
  ggtitle('Synthetic Controls for Medicaid Expansion') +
  xlab('Time') +
  ylab('Uninsured Rate')
```
```{r}
#Visualizing the ATE on states that implemented Medicaid work requirements.
ppool_syn_summ$att %>%
  filter (Level %in% c("Kentucky", "Indiana", "Arkansas" , "New Hampshire" , "Arizona" , "Michigan", "Ohio" ,
                       "Utah", "Wisconsin"))%>%
  ggplot(aes(x = Time, y = Estimate, color = Level)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = "bottom") +
  ggtitle('Synthetic Controls for Medicaid Expansion') +
  xlab('Time') +
  ylab('Uninsured Rate')
```

- Estimate a multisynth model using time cohorts. For the purpose of this exercise, you can simplify the treatment time so that states that adopted Medicaid expansion within the same year (i.e. all states that adopted epxansion in 2016) count for the same cohort. Plot the treatment effects for these time cohorts.


```{r}

#multisynth model time cohorts

ppool_syn_time <- multisynth(uninsured_rate ~ trt, State, year, 
                        medicaid_expansion, n_leads = 10 , time_cohort = TRUE)

ppool_syn_time_summ <- summary(ppool_syn_time)

ppool_syn_time_summ

```

```{r}
ppool_syn_time_summ$att %>%
  ggplot(aes(x = Time, y = Estimate, color = Level)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = 'None') +
  ggtitle('Synthetic Controls for Medicaid Expansion') +
  xlab('Time') +
  ylab('Uninsured Rate') +
  facet_wrap(~Level)
```
```{r}
#Augmented synthetic control model

scm_gsyn <- multisynth(uninsured_rate ~ trt, State, year, 
                        medicaid_expansion, n_leads = 10, 
                        fixedeff = T, n_factors = 2)

scm_gsyn_summ <- summary(scm_gsyn)
```

```{r}
#Visualizing augmented synthetic control model

scm_gsyn_summ$att %>%
  ggplot(aes(x = Time, y = Estimate, color = Level)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0) +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = 'None') +
  ggtitle('Augmented Synthetic Controls for Medicaid Expansion') +
  xlab('Time') +
  ylab('Uninsured Rate') +
  facet_wrap(~Level)
```

## Discussion Questions

- One feature of Medicaid is that it is jointly administered by the federal government and the states, and states have some flexibility in how they implement Medicaid. For example, during the Trump administration, several states applied for waivers where they could add work requirements to the eligibility standards (i.e. an individual needed to work for 80 hours/month to qualify for Medicaid). Given these differences, do you see evidence for the idea that different states had different treatment effect sizes?
- **Answer**: There does appear to be evidence that different states had different treatment effect sizes. We see this variation in several of our plots. For some states, like California and Arizona, there appear to be larger decreases in uninsured rates. However, for some states like Hawaii and Vermont, we see that uninsured rates increase. We also see variation in treatment effects for states that implemented Medicaid work requirements, though for this subset of states we do see a general decrease in uninsured rates.

- Do you see evidence for the idea that early adopters of Medicaid expansion enjoyed a larger decrease in the uninsured population?
- **Answer**: We do see evidence that early adopters--states that implemented a Medicaid expansion between 2014 and 2016--experienced a larger decrease in their uninsured population. Using a multisynth model with time cohort, we see that states that implemented the expansion in 2015 and 2016 appear to have experienced the strongest treatment effects. 

# General Discussion Questions

- Why are DiD and  synthetic control estimates well suited to studies of aggregated units like cities, states, countries, etc?
- **Answer**: Both DiD and synthetic control estimates are well suited to studies of aggregated units because they are able to form comparable counterfactuals to the treatment group that allows us to measure causal effects, albeit in different ways. 

DiD is useful for aggregated data because it attempts to measure the effect of an exogenous shock (e.g. policy intervention) on the affected group and is not necessarily concerned with variation (e.g., randomization is not possible) at the individual level. In the absence of randomization of treatment, a group of untreated units (for which observations over time exist) can serve as the counterfactual to the treated group, or what the treated group would have looked like in the absence of treatment, thereby producing an estimate for the treatment effect. When the parallel trends assumption is met, DiD can subtract out the effect of confounders over time from both the treatment and control groups. For example, in assessing the effect of the passage of a health insurance policy on states, DiD method accounts for, and is able to take care of, confounders that might have produced differences in outcomes between the treated and untreated states, as long as states are assumed to have changed in the same manner over time (parallel trends assumption). 

Similarly, synthetic control estimates are also good for studying aggregated units because aggregate data aids in forming the appropriate counterfactual to the treatment group. When it is difficult to find one single unit to which to compare the treatment group, synthetic control uses a combination of multiple differently-weighted control units to serve as the counterfactual. 


- What role does selection into treatment play in DiD/synthetic control versus regression discontinuity? When would we want to use either method?
- **Answer**: Both make different assumptions about the presence of unobserved confounders. With DiD, it is assumed that there may be unobserved confounders with the selection process (Abadie, Diamond, and Hainmueller 495), but because of the DiD design, variation due to those confounders is eliminated by the choice of comparative case and the difference over time. The risk lies in the potential to choose a treatment and comparison that do not have parallel trends, meaning that variation is not adequately eliminated. 

Regression discontinuity is different in that local regressions are run on either side of the treatment threshold (Cortes, Goodman, and Nomi 117). In contrast to DiD then, it is very important that the assumption of unobserved confounders is not violated, and that units very near the threshold are very similar in their covariates. If there is evidence of sorting at the threshold, regression discontinuity estimates of the local average treatment effect are biased. 

We think regression discontinuity makes more sense in the research designs with many units, a variable that can be used as a running variable (meaning it is continuous and independent from treatment assignment), and to answer a question that is not too concerned with treatment effects far away from the threshold (i.e., outside of the bandwidth). DiD/synthetic control makes more sense when comparing fewer units and when the research question is concerned with the comparison of trends that stretch farther before and after the treatment threshold.
